---
title: "Exploring the Iris Dataset with R"
author: "Wahid Raziqi"
date: today
categories: [r, data-analysis]
---

# Introduction

The Iris dataset is a classic dataset used in statistics and data analysis. It contains measurements of iris flowers from three species:

- Setosa
- Versicolor
- Virginica

The dataset is included in R, making it convenient for a reproducible analysis.

# Load packages and data

```{r}
#| message: false
#| warning: false

library(dplyr)
library(ggplot2)

data(iris)
```

```{r}
iris_summary <- iris |>
  group_by(Species) |>
  summarise(
    mean_sepal_length = mean(Sepal.Length),
    mean_sepal_width = mean(Sepal.Width),
    mean_petal_length = mean(Petal.Length),
    mean_petal_width = mean(Petal.Width)
  )

iris_summary
```
# Summary statistics

I calculated the mean sepal length, sepal width, petal length, and petal width for each iris species.

```{r}
#| label: fig-iris-measurements
#| fig-cap: "Sepal length versus petal length for the three iris species."

ggplot(
  iris,
  aes(
    x = Sepal.Length,
    y = Petal.Length,
    colour = Species
  )
) +
  geom_point() +
  labs(
    title = "Iris species comparison",
    x = "Sepal length (cm)",
    y = "Petal length (cm)",
    colour = "Species"
  ) +
  theme_minimal()
```

The visualization shows a clear separation between Setosa and the other two species based on petal length. Versicolor and Virginica have more similar measurements and show some overlap.

# Discussion

This analysis demonstrates how R, dplyr, and ggplot2 can be used together in a Quarto document. The calculations and figure are generated when the document is rendered, making the analysis reproducible.

# Data source

Data: Iris dataset, included with R as the built-in `iris` dataset. The dataset is documented in the R documentation for the `datasets` package.