---
title: "R and Python in One Post"
author: "Wahid Raziqi"
date: today
categories: [r, python, reticulate]
---

# Introduction

This post demonstrates how R and Python can be used together in the same Quarto document. The `reticulate` package allows an object created in R to be used in Python, and a result calculated in Python can then be returned to R.



```{r}
library(reticulate)

r_values <- c(10, 20, 30, 40, 50)

r_values
```

# Use the R object in Python

```{python}
python_values = r.r_values

print("Values received from R:")
print(python_values)

python_mean = sum(python_values) / len(python_values)

print("Mean calculated in Python:")
print(python_mean)
```

# Use the Python result in R

```{r}
final_mean <- py$python_mean

cat("The mean calculated in Python and returned to R is:", final_mean)
```

# Conclusion

This example demonstrates how `reticulate` allows R and Python to work together in the same Quarto document. The values were created in R, the mean was calculated in Python, and the Python result was passed back to R.